\Large \textbf{ Phenotype and gene ontology enrichment as guides for disease modeling in \cel{} }

### 2017-02-03 pea_paper_docs/tex/pea.tex
wrote a bit for the manuscript

\newcommand{\qval}[1]{
                      \ensuremath{
                                  q<10^{-#1}
                                  }
                      }

@@

\section*{Introduction}
The last decade has seen an explosion of techniques capable of genome-wide
measurements. Some examples of genome-wide tools include
RNA-seq~\cite{Mortazavi2008} to measure gene expression or
CHIP-seq~\cite{Johnson2007} to measure protein binding. These tools are capable
of generating large quantities of data. Understanding these data, and generating
hypotheses from them remains challenging. A common approach used to understand
these datasets is to reduce the dimensionality of the data via enrichment
analyses of ontologies~\cite{TheGeneOntologyConsortium2000a}, which helps
researchers understand what terms are
overrepresented beyond random levels. By analyzing overrepresented terms in
aggregate, researchers can better understand what biological processes were most
affected in a given experiment, and form hypotheses about what is
happening~\cite{Rhee2008}.
This approach is limited by what ontologies can be tested for enrichment.
The best-known ontology for biological research is the Gene Ontology (GO), which
provides a controlled language to describe molecular and cellular functions of
genes~\cite{TheGeneOntologyConsortium2000a}.
In \cel{}, gene, tissue and phenotype ontologies exist
% which provide
% controlled languages
with which to describe \cel{} anatomy and phenotypes
respectively~\cite{Schindelman2011,Lee2003}. These ontologies are curated by
professional curators at WormBase, which is a repository of all \cel{}
data~\cite{Howe2016}.
However, enrichment tools
only exist for gene and tissue ontologies
in the community today (see for example~\cite{Chikina2009,Mi2013,Angeles-Albores2016}).
Another limitation is that
tissue enrichment testing is not offered on the same websites as GO enrichment
testing, which requires users to test their data on different websites that may
or may not use different methodologies to detect enrichment.

Another way to use enrichment tools is for evolutionary comparison purposes.
In molecular biology it is often useful to know when a gene is homologous
between two species---that is to say, common by descent---because knowledge of
homology often brings with it knowledge of function. Indeed, many
important gene regulatory networks (GRNs) are conserved between organisms as
highly diverged as nematodes and humans (for example, see~\cite{Sternberg1998}).
While genes
and GRNs may be conserved between species, their outputs often differ however.
For example, although homeobox genes are known to be important for limb
formation in many animals~\cite{Zakany2007}, these genes do not form limbs in nematodes.
The concept of a phenolog has been put forward to explain relationships between
phenotypes that have the same underlying genetic
regulatory network~\cite{McGary2010,Lehner2013}.
Formally, two phenotypes are phenologs of each other if the homologs of the
genes that cause a phenotype in an organism cause a second phenotype in another.

To study a clinically relevant disease in a non-human, an appropriate
model has to be established. A straightforward method towards establishing a
disease model in
\cel{} is to link a disease to a causal gene, then to identify the homologous
gene in \cel{} and then to study the function of the genetic homolog to
extrapolate back to humans. However, this method relies on the existence of
known disease genes and requires that the homolog have a phenotype that can be
reliably identified and studied. A fundamentally different way to establish a
disease model in \cel{} would be to identify the phenologs of the disease to be
studied in \cel{} by identifying disease-associated human genes in an unbiased
manner through genome-wide association studies (GWAS) and identified candidate
homolog genes in \cel{}. The homologs can be used to identify \cel{} disease
phenologs, which can in turn be used as the basis for screens to identify
genes that are associated with that phenolog. Approaches similar to this have
been successfully used in the past to make non-obvious links between phenotypes
in different species~\cite{McGary2010}.

@@

An additional problem with genome-wide
queries of \cel{} states (be they developmental, such as
L1, L2, dauer; behavioral states such as awake versus asleep; or other) is that
they do not
always have a straightforward interpretation in terms of phenotypes. In these
situations, researchers must rely on intuition to select a phenotype to screen
for. As a result, many hits may go unexplored that would prove fruitful. The
question of how to design a screen that is maximally informative is an important
question that has so far not been addressed within this community.

@@

\subsection*{Human disease phenolog identification}
We used the GWAS EBI-NHGRI catalog~\cite{MacArthur2016} to extract information
on all genome-wide
association studies deposited there. We only selected traits that had $>300$
associated genes. We identified 24 traits that met our criteria. Next, we used
DIOPT~\cite{Hu2011} to identify candidate homologs for the genes associated with these
traits. Briefly, DIOPT combines a large number of methods for identifying homologs
and returns homolog candidates associated with a compound score. Depending on the score,
homologs can be considered `high', `moderate' or `low' rank, reflecting confidence
in the homology. Many-to-one and one-to-many homology relationships are allowed
in DIOPT, reflecting a mixture of uncertainty and family expansion/reduction.
For our study, we only accepted homolog candidates with `high' or `moderate' scores
and we did not insist on a one-to-one relationship between genes.

\begin{figure}[htbp]
  \renewcommand{\familydefault}{\sfdefault}\normalfont{}
  \centering
  \includegraphics[width=\linewidth]{gwas-design.pdf}
  \caption{Experimental design for human-nematode phenologue identification.
           We used GWAS candidates from the EBI-NHGRI catalog to identify
           disease associated genes, then used DIOPT to identify candidate
           homologs in \cel{}. Homologs were used to run phenotype, gene
           and tissue ontology enrichment analyses to identify disease
           phenologues.}
\label{fig:gwas}
\end{figure}

After we identified worm homologs for each trait, we reassessed how many traits
still had $>100$ gene candidates, and dropped all traits that had less than this
for our analysis. We identified 18 traits that met this criteria. The gene
lists for each of these 18 traits were then analyzed for gene, tissue and
phenotype enrichment (see~Fig.\ref{fig:gwas}). Tissue enrichment was
performed using the WormBase Tissue Enrichment Analysis (TEA)
tool~\cite{Angeles-Albores2016}.

@@

\subsection*{Applying the WormBase enrichment suite}
To discover phenologs, we first needed to identify genes that
contribute to a disease in an unbiased manner. One way to discover gene
associations in an unbiased manner is to perform GWSA in human populations.
Therefore, we used the GWAS NHGRI-EBI
Catalog~\cite{MacArthur2016} to identify genes associated with human diseases. We found the
best nematode candidate homologs for these genes using DIOPT~\cite{Hu2011} and applied
our enrichment suite to each of these gene regulatory networks.

\subsubsection*{Obesity-related traits}
Obesity-related traits is a category within the GWAS NHGRI-EBI catalog that
pools studies that have measured obesity and other traits associated with
obesity, such as heart rate, physical activity, hormone levels, body composition
and cholesterol levels. Since this category includes many parameters,
we expected there would be many phenologs. GWAS studies have identified
\hobesity{} genes associated with these traits. Using DIOPT, we found
\wobesity{} homologs for these genes. In total, $341/\wobesity{}$ genes had
at least one phenotype annotation; $548/\wobesity{}$ had at least one gene ontology term
annotation; and $427/\wobesity{}$ had at least one tissue term annotation.

@@

Lupus-associated homologs were reasonably well annotated. Slightly more than half
of the genes had at least one phenotype annotation ($76/\wlupus{}$) and almost
all genes were annotated to at least one tissue or gene ontology term
($104/\wlupus{}$ and $115/\wlupus$ genes respectively).
We found that Lupus-associated homologs were enriched in `aneuploidy' (7 genes,
\qval{1}) and `meiotic chromosome segregation' (8 genes, \qval{1}). `Cell fate
transformation' (6 genes, \qval{1}), and `excess intestinal cells'
(5 genes, \qval{1}) were also overrepresented, as was `male tail morphology'
(6 genes, \qval{1}). Finally, the phenotype `nonsense mRNA accumulation' was also
enriched (5 genes, \qval{1}) (see Fig.~\ref{fig:lupus}).

@@

% TODO: Is no tissue enrichment observed bc poor annotations, or because
% all tissues randomly represented?
The only phenotype that was enriched for these homologs was `short' (10 genes,
\qval{4}), even though 64 homologs were associated with at least
one phenotype term. No tissue was enriched in this dataset. Because
82 genes are annotated to have expression in at least one tissue, the lack of
enrichment does not reflect ignorance about the sites of expression of these
genes. GEA showed that enriched
molecular functions for these genes include `collagen trimer' (22 genes,
\qval{15}). However, this term was enriched as the result of degeneracy in the
homolog candidates for the SFTPD gene. Other terms included
 `glycosylation' (10 genes, \qval{4}) and `Golgi apparatus' (11
genes, \qval{3}), but these terms were enriched as the result of degenerate homolog
candidates for the human gene B3GNT7 which encodes a
beta-1,3-N-acetylgalactosaminyltransferase.

The `short' phenotype was the result of the \emph{cat-4}, \emph{dpy-7},
\emph{rnt-1}, \emph{sem-4}, \emph{unc-116}, \emph{ocrl-1} and some genes in the
\emph{fat} family. Although these genes are bound by a common phenotype, any
genetic relationships between these genes are not immediately clear. Some genes,
like \emph{sem-4} and \emph{rnt-1} are likely transcription factors with roles
in development (including hypodermal development)~\cite{Desai1988,Ji2004}.
Others are molecular motors (\emph{unc-116}) that are broadly expressed throughout
the body of \cel{}. Yet others have known roles in neuron and muscle function,
such as \emph{ocrl-1} and \emph{cat-4}. The `short' phenotype is a subset of
the `body length variant' phenotype. Body length in \cel{} can be controlled via
cell size, shape and number\cite{}; alternatively, cuticle development can alter body
shape~\cite{}; finally, muscles can alter the effective body length artificially~\cite{}.

@@

\subsection*{Ontology Enrichment as an aid for screen design}
An additional use for a tool like PEA would be as a tool to help guide and
design screens to identify genes from an RNA-seq or other genome-wide experiment
for further study. This would be particularly useful in cases when researchers
may not know what phenotype to expect, in which case PEA can guide selection of
a phenotype. Another use case is a scenario where the phenotype under study is
not easy to screen for. By finding phenologs to the phenotype of interest, the
researcher can design an easier screen for genes that affect the phenolog in
question, then re-test genes for the original phenotype of interest.

@@

In addition, TEA points at the \cel{} gonad primordium, the somatic gonad
and early embryonic cells as the sites where genes associated with ciliary neurons
are enriched. The `male distal tip cell' is a tissue
that is overrepresented in this dataset, but `distal tip cell' is not enriched,
which suggests that structures that are present only in the male tissue are
overrepresented in this dataset.

@@

\begin{figure}[htbp]
  \renewcommand{\familydefault}{\sfdefault}\normalfont{}
  \centering
  \includegraphics[width=\linewidth]{Egl-phenotype.pdf}
  \caption{The Egl phenotype is a complex phenotype that is the result of
  interactions between many tissues. To dissect the contributions of individual
  tissues to generating an Egl phenotype, we obtained all genes annotated with
  with an Egl term. \textbf{A} We used TEA on the set of Egl genes to identify
  enriched tissues. \textbf{B} Anatomic diagram showing the tissues that are
  most enriched in the Egl gene set. Color coding shows the qualitative
  ordering of enrichment (red-Most enriched, yellow-least enriched). For clarity,
  not all cells are shown. All vm1 (4 cells) and vm2 (4 cells) muscles are
  symmetrical arranged around the vulva, but only 2 cells are shown for each.
  There are two seam cells on the left and right side of the vulva, but only
  cell on the right is shown. Only cells in the adult are shown.}
\label{fig:egl}
\end{figure}

@@

We calculated the conditional probability that a gene has an Egl phenotype given
that it's expression pattern includes a tissue term $X$ and we searched for the
tissue terms that maximized this probability. The list of terms that maximized this
probability reflected the results from running TEA on the subset of genes that have
an Egl phenotype. We also calculated the conditional probability that a gene has
expression in a tissue term $X$ given that it is annotated with an Egl phenotype
and we searched for terms that maximized this probability. The terms that maximized
this probability were `nervous system', `pharynx' (a body part with a lot of neurons),
`sex organ' and `tail' (a body part with neurons and hypodermis). In general, the
terms that had a high $P(\text{Egl}|X)$ did not have a high $P(X|\text{Egl})$.
Additionally, the terms that had a high $P(X|\text{Egl})$ are broad terms that
include a lot of cells, whereas the terms that had a high $P(\text{Egl}|X)$
were considerably more specific.
We conclude that the Egl phenotype arises from a small set of tissues. The Egl
phenotype can be best predicted by genes with expression patterns that include
at least one of a small number of cells (mainly vul cells, HSN). On the other
hand, answering whether the expression pattern of a gene includes a particular
anatomic region or tissue given that the gene has an Egl phenotype is hard to
do for small tissues or single cells. However, guesses about what functional system
or broad anatomic region an Egl gene is expressed in can be answered with
confidence ($\sim70\%$ of the time, an Egl mutant is expressed in
the nervous system).

>>>>

\newcommand{\qval}[1]{\ensuremath{q<10^{-#1}}}

@@

\section*{Introduction}
The last decade has seen an explosion of techniques capable of genome-wide
measurements. Some examples of genome-wide tools include
RNA-seq~\cite{Mortazavi2008} to measure gene expression or
CHIP-seq~\cite{Johnson2007} to measure protein binding to chromatin. These tools are capable
of generating large quantities of data. Understanding these data, and generating
hypotheses from them remains challenging. A common approach used to understand
these datasets is to reduce the dimensionality of the data via enrichment
analyses of ontologies~\cite{TheGeneOntologyConsortium2000a}, which helps
researchers understand what terms are
overrepresented beyond random levels. By analyzing overrepresented terms in
aggregate, researchers can better understand what biological processes were most
affected in a given experiment, and form hypotheses about what is
happening~\cite{Rhee2008}.
This approach is limited by what ontologies can be tested for enrichment.
The best-known ontology for biological research is the Gene Ontology (GO), which
provides a controlled language to describe molecular and cellular functions of
genes~\cite{TheGeneOntologyConsortium2000a}.
In \cel{}, gene, tissue and phenotype ontologies exist
% which provide
% controlled languages
with which to describe \cel{} anatomy and phenotypes
respectively~\cite{Schindelman2011,Lee2003}. These ontologies are curated by
professional curators at WormBase, which is a repository of all \cel{}
data~\cite{Howe2016}.
However, enrichment tools
only exist for gene and tissue ontologies
in the community today (see for example~\cite{Chikina2009,Mi2013,Angeles-Albores2016}).
Another limitation is that
tissue enrichment testing is not offered on the same websites as GO enrichment
testing, which requires users to test their data on different websites that may
or may not use different methodologies to detect enrichment.

Another way to use enrichment tools is for evolutionary comparison purposes.
In molecular biology it is often useful to know when a gene is homologous
between two species---that is to say, common by descent---because knowledge of
homology often brings with it knowledge of function. Indeed, many
important gene regulatory networks (GRNs) are conserved between organisms as
highly diverged as nematodes and humans (for example, see~\cite{Sternberg1998}).
While genes and GRNs may be conserved between species, their outputs often differ.
For example, the gene Pax6 (Eyeless in \emph{Drosophila melanogaster})
is involved in eye formation in humans and fruit-flies~\cite{Quiring1994}. Although
nematodes have conserved this gene, they do not have eyes~\cite{Zhang1995,Chisholm1995}.
The concept of a phenolog has been put forward to explain relationships between
phenotypes that have the same underlying genetic
regulatory network~\cite{McGary2010,Lehner2013}.
Formally, two phenotypes are phenologs of each other if the orthologs of the
genes that cause a phenotype in an organism cause a second phenotype in another.

To study a clinically relevant disease in a non-human, an appropriate
model has to be established. A straightforward method towards establishing a
disease model in
\cel{} is to link a disease to a causal gene, then to identify the homologous
gene in \cel{} and then to study the function of the genetic homolog to
extrapolate back to humans. However, this method relies on the existence of
known disease genes and requires that the homolog have a phenotype that can be
reliably identified and studied. A fundamentally different way to establish a
disease model in \cel{} would be to identify the phenologs of the disease to be
studied in \cel{} by identifying disease-associated human genes in an unbiased
manner through genome-wide association studies (GWAS) and identified candidate
homolog genes in \cel{}. The orthologs can be used to identify \cel{} disease
phenologs, which can in turn be used as the basis for screens to identify
genes that are associated with that phenolog. Approaches similar to this have
been successfully used in the past to make non-obvious links between phenotypes
in different species~\cite{McGary2010}.

@@

An additional problem with genome-wide
queries of \cel{} states (be they developmental, such as
L1, L2, dauer; behavioral states such as awake versus asleep; or other) is that
they do not
always have a straightforward interpretation in terms of phenotypes. In these
situations, researchers must rely on intuition to select a phenotype for which to
screen. As a result, many hits may go unexplored that would prove fruitful. The
question of how to design a screen that is maximally informative is an important
question that has so far not been addressed within this community.

@@

\subsection*{Human disease phenolog identification}
We used the GWAS EBI-NHGRI catalog~\cite{MacArthur2016} to extract information
on all genome-wide
association studies deposited there. We only selected traits that had $>300$
associated genes. We identified 24 traits that met our criteria. Next, we used
DIOPT~\cite{Hu2011} to identify candidate orthologs for the genes associated with these
traits. Briefly, DIOPT combines a large number of methods for identifying orthologs
and returns homolog candidates associated with a compound score. Depending on the score,
orthologs can be considered `high', `moderate' or `low' rank, reflecting confidence
in the homology. Many-to-one and one-to-many homology relationships are allowed
in DIOPT, reflecting a mixture of uncertainty and family expansion/reduction.
For our study, we only accepted homolog candidates with `high' or `moderate' scores
and we did not insist on a one-to-one relationship between genes.

\begin{figure}[htbp]
  \renewcommand{\familydefault}{\sfdefault}\normalfont{}
  \centering
  \includegraphics[width=\linewidth]{gwas-design.pdf}
  \caption{Experimental design for human-nematode phenologue identification.
           We used GWAS candidates from the EBI-NHGRI catalog to identify
           disease associated genes, then used DIOPT to identify candidate
           orthologs in \cel{}. Orthologs were used to run phenotype, gene
           and tissue ontology enrichment analyses to identify disease
           phenologues.}
\label{fig:gwas}
\end{figure}

After we identified worm orthologs for each trait, we reassessed how many traits
still had $>100$ gene candidates, and dropped all traits that had less than this
for our analysis. We identified 18 traits that met this criteria. The gene
lists for each of these 18 traits were then analyzed for gene, tissue and
phenotype enrichment (see~Fig.\ref{fig:gwas}). Tissue enrichment was
performed using the WormBase Tissue Enrichment Analysis (TEA)
tool~\cite{Angeles-Albores2016}.

@@

\subsection*{Applying the WormBase enrichment suite}
To discover phenologs, we first needed to identify genes that
contribute to a disease in an unbiased manner. One way to discover gene
associations in an unbiased manner is to perform GWSA in human populations.
Therefore, we used the GWAS NHGRI-EBI
Catalog~\cite{MacArthur2016} to identify genes associated with human diseases. We found the
best nematode candidate orthologs for these genes using DIOPT~\cite{Hu2011} and applied
our enrichment suite to each of these gene regulatory networks.

@@

Lupus-associated orthologs were reasonably well annotated. Slightly more than half
of the genes had at least one phenotype annotation ($76/\wlupus{}$) and almost
all genes were annotated to at least one tissue or gene ontology term
($104/\wlupus{}$ and $115/\wlupus$ genes respectively).
We found that Lupus-associated orthologs were enriched in `aneuploidy' (7 genes,
\qval{1}) and `meiotic chromosome segregation' (8 genes, \qval{1}). `Cell fate
transformation' (6 genes, \qval{1}), and `excess intestinal cells'
(5 genes, \qval{1}) were also overrepresented, as was `male tail morphology'
(6 genes, \qval{1}). Finally, the phenotype `nonsense mRNA accumulation' was also
enriched (5 genes, \qval{1}) (see Fig.~\ref{fig:lupus}).

@@

% TODO: Is no tissue enrichment observed bc poor annotations, or because
% all tissues randomly represented?
The only phenotype that was enriched for these orthologs was `short' (10 genes,
\qval{4}), even though 64 orthologs were associated with at least
one phenotype term. No tissue was enriched in this dataset. Because
82 genes are annotated to have expression in at least one tissue, the lack of
enrichment does not reflect ignorance about the sites of expression of these
genes. GEA showed that enriched
molecular functions for these genes include `collagen trimer' (22 genes,
\qval{15}). However, this term was enriched as the result of degeneracy in the
homolog candidates for the SFTPD gene. Other terms included
 `glycosylation' (10 genes, \qval{4}) and `Golgi apparatus' (11
genes, \qval{3}), but these terms were enriched as the result of degenerate homolog
candidates for the human gene B3GNT7 which encodes a
beta-1,3-N-acetylgalactosaminyltransferase.

The `short' phenotype was the result of the \emph{cat-4}, \emph{dpy-7},
\emph{rnt-1}, \emph{sem-4}, \emph{unc-116}, \emph{ocrl-1} and some genes in the
\emph{fat} family. Although these genes are bound by a common phenotype, any
genetic relationships between these genes are not immediately clear. Some genes,
like \emph{sem-4} and \emph{rnt-1} are likely transcription factors with roles
in development (including hypodermal development)~\cite{Desai1988,Ji2004}.
Others are molecular motors (\emph{unc-116}) that are broadly expressed throughout
the body of \cel{}. Yet others have known roles in neuron and muscle function,
such as \emph{ocrl-1} and \emph{cat-4}. The `short' phenotype is a subset of
the `body length variant' phenotype. Body length in \cel{} can be controlled via
cell size or shape\cite{}; alternatively, cuticle development can alter body
shape~\cite{}; finally, muscles can alter the effective body length
due to their contraction state~\cite{}.

@@

\subsubsection*{Obesity-related traits}
Obesity-related traits is a category within the GWAS NHGRI-EBI catalog that
pools studies that have measured obesity and other traits associated with
obesity, such as heart rate, physical activity, hormone levels, body composition
and cholesterol levels. Since this category includes many parameters,
we expected there would be many phenologs. GWAS studies have identified
\hobesity{} genes associated with these traits. Using DIOPT, we found
\wobesity{} orthologs for these genes. In total, $341/\wobesity{}$ genes had
at least one phenotype annotation; $548/\wobesity{}$ had at least one gene ontology term
annotation; and $427/\wobesity{}$ had at least one tissue term annotation.

\subsection*{Ontology enrichment as an aid for screen design}
An additional use for a tool like PEA would be as a tool to help guide and
design screens to identify genes from an RNA-seq or other genome-wide experiment
for further study. This would be particularly useful in cases when researchers
may not know what phenotype to expect, in which case PEA can guide selection of
a phenotype. Another use case is a scenario where the phenotype under study is
not easy to screen for. By finding phenologs to the phenotype of interest, the
researcher can design an easier screen for genes that affect the phenolog in
question, then re-test genes for the original phenotype of interest.

@@

In addition, TEA points at the \cel{} gonad primordium, the somatic gonad
and early embryonic cells as the sites where genes associated with ciliary neurons
are enriched. The `male distal tip cell' is a tissue
that is overrepresented in this dataset, but `distal tip cell' is not enriched.
In \cel{} the hermaphrodite distal tip cells (DTCs) and the male DTC are very
similar to each other in their biological functions (both maintain a stem cell
niche in the distal gonad). However, the male DTC is non-migratory, whereas the
hermaphrodite DTCs are migratory. Therefore, the term `male distal tip cell'
may reflect cellular aspects that are correlated with the non-migratory aspects
of the DTC biology, such as maintenance of proliferation.

@@

\begin{figure}[htbp]
  \renewcommand{\familydefault}{\sfdefault}\normalfont{}
  \centering
  \includegraphics[width=\linewidth]{Egl-phenotype.pdf}
  \caption{The Egl phenotype is a complex phenotype that is the result of
  interactions between many tissues. To dissect the contributions of individual
  tissues to generating an Egl phenotype, we obtained all genes annotated with
  with an Egl term. \textbf{A} We used TEA on the set of Egl genes to identify
  enriched tissues. \textbf{B} Anatomic diagram showing the tissues that are
  most enriched in the Egl gene set. Color coding shows the qualitative
  ordering of enrichment (red-Most enriched, yellow-least enriched). For clarity,
  not all cells are shown. All vm1 (4 cells) and vm2 (4 cells) muscles are
  symmetrical arranged around the vulva, but only 2 cells are shown for each.
  There are two seam cells on the left and right side of the vulva, but only
  cell on the right is shown. Only terminally differentiated cells are shown.}
\label{fig:egl}
\end{figure}

@@

We calculated the conditional probability that a gene has an Egl phenotype given
that it's expression pattern includes a tissue term $X$ and we searched for the
tissue terms that maximized this probability. The list of terms that maximized this
probability reflected the results from running TEA on the subset of genes that have
an Egl phenotype. We also calculated the conditional probability that a gene has
expression in a tissue term $X$ given that it is annotated with an Egl phenotype
and we searched for terms that maximized this probability (see
Table~\ref{tab:cond_probs}). The terms that maximized this probability were
`nervous system', `pharynx' (a body part with a lot of neurons),
`sex organ' and `tail' (a body part with neurons and hypodermis). In general, the
terms that had a high $P(\text{Egl}|X)$ did not have a high $P(X|\text{Egl})$.
Additionally, the terms that had a high $P(X|\text{Egl})$ are broad terms that
include a lot of cells, whereas the terms that had a high $P(\text{Egl}|X)$
were considerably more specific.
We conclude that the Egl phenotype arises from a small set of tissues. The Egl
phenotype can be best predicted by genes with expression patterns that include
at least one of a small number of cells (mainly vul cells, HSN). On the other
hand, answering whether the expression pattern of a gene includes a particular
anatomic region or tissue given that the gene has an Egl phenotype is hard to
do for small tissues or single cells. However, guesses about what functional system
or broad anatomic region an Egl gene is expressed in can be answered with
confidence ($\sim70\%$ of the time, an Egl mutant is expressed in
the nervous system).
